---
title: "FORMATIVE ASSESSMENT 5"
author: "Lindsy Rossel C. Masicat"
date: "2024-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(ggplot2)
```

## Problem 1
### 8.18. List all samples of size n ¼ 2 that are possible (with replacement) from the population in Problem 8.17. Using R, plot the sampling distribution of the mean to show that \[ \mu_{\bar{x}} = \mu \], and show that \[ \sigma^2_{\bar{x}} = \frac{\sigma^2}{2} \]
.
```{r samples, echo=FALSE}
population <- c(9, 12, 15)
probabilities <- rep(1/3, 3) 

samples <- expand.grid(population, population)
colnames(samples) <- c("X1", "X2")


samples$mean <- rowMeans(samples)
samples$p_xbar <- rep(1/9, nrow(samples))

samples$xbar_p_xbar <- samples$mean * samples$p_xbar
samples$xbar2_p_xbar <- (samples$mean^2) * samples$p_xbar

final_table <- data.frame(
  X1 = samples$X1,
  X2 = samples$X2,
  mean = samples$mean,
  xbar = samples$mean,
  `p(xbar)` = samples$p_xbar,
  `xbar * p(xbar)` = samples$xbar_p_xbar,
  `xbar^2 * p(xbar)` = samples$xbar2_p_xbar
)

knitr::kable(final_table, 
             col.names = c("X1", "X2", "mean", "xbar", "p(xbar)", "xbar * p(xbar)", "xbar^2 * p(xbar)"), 
             caption = "Sampling Distribution Table")

sampling_dist <- data.frame(mean = samples$mean)

ggplot(sampling_dist, aes(x = mean)) +
  geom_bar(stat = "count", fill = "steelblue") +
  labs(title = "Sampling Distribution of the Mean",
       x = "Sample Mean",
       y = "Frequency")

```

## Problems 2
### 8.21. A population consists of the four numbers 3, 7, 11, and 15. Consider all possible samples of size 2 that can be drawn with replacement from this population. Find (a) the population mean,(b) the population standard deviation, (c) the mean of the sampling distribution of means, and (d ) the standard deviation of the sampling distribution of means. Verify parts (c) and (d ) directly from (a) and (b) by using suitable formulas.

### Population Mean and Standard Deviation

The population consists of the numbers 3, 7, 11, and 15.

### (a) Population Mean \( \mu \)
The population mean is calculated as:
$$ \mu = \frac{3 + 7 + 11 + 15}{4} = \frac{36}{4} = 9 $$

### (b) Population Standard Deviation \( \sigma \)
The population standard deviation is calculated as:
$$\sigma = \sqrt{\frac{(3 - 9)^2 + (7 - 9)^2 + (11 - 9)^2 + (15 - 9)^2}{4}}$$
First, compute the squared differences:
$$(3 - 9)^2 = 36, \quad (7 - 9)^2 = 4, \quad (11 - 9)^2 = 4, \quad (15 - 9)^2 = 36$$
Summing these values gives:
$$36 + 4 + 4 + 36 = 80$$
Now, divide by the number of elements:
$$\sigma = \sqrt{\frac{80}{4}} = \sqrt{20} = 4.47$$

### (c) Mean of the Sampling Distribution of Means \( \mu_{\bar{X}} \)

Since we are sampling with replacement, the mean of the sampling distribution of means is equal to the population mean:
$$\mu_{\bar{X}} = \mu = 9$$

### (d) Standard Deviation of the Sampling Distribution of Means \( \sigma_{\bar{X}} \)

The standard deviation of the sampling distribution of means is calculated as:
$$\sigma_{\bar{X}} = \frac{\sigma}{\sqrt{n}} = \frac{4.47}{\sqrt{2}} = \frac{4.47}{1.414} = 3.16$$

### Verification:
The calculations confirm that:

1. $$ \mu_{\bar{X}} = \mu = 9 $$
2. $$\sigma_{\bar{X}} = \frac{\sigma}{\sqrt{n}} = 3.16 $$


### 8.34. Find the probability that of the next 200 children born, (a) less than 40% will be boys, (b) between 43% and 57% will be girls, and (c) more than 54% will be boys. Assume equal probabilities for the births of boys and girls.

We are given that the probability of a child being a boy is \( p = 0.5 \), and the number of births is \( n = 200 \). The number of boys \( X \) follows a binomial distribution:

$$X \sim \text{Binomial}(n = 200, p = 0.5)$$

The binomial distribution can be approximated by a normal distribution:

$$X \sim N(\mu = np, \sigma^2 = np(1-p))$$

The mean and standard deviation are:

$$\mu = np = 200 \times 0.5 = 100$$
$$\sigma^2 = np(1 - p) = 200 \times 0.5 \times 0.5 = 50$$
$$\sigma = \sqrt{50} \approx 7.071$$

---

### (a) Probability that less than 40% will be boys

We want to compute \( P(X < 0.40 \times 200) \), which translates to:

$$P(X < 80)$$

To approximate this using the normal distribution, we convert to the z-score:

$$Z = \frac{X - \mu}{\sigma} = \frac{80 - 100}{7.071} = \frac{-20}{7.071} \approx -2.828$$

Now, using the cumulative distribution function of the standard normal distribution:

$$P(Z < -2.828) \approx 0.0023$$

Thus, the probability that less than 40% will be boys is approximately:

$$P(X < 80) \approx 0.0023$$

---

### (b) Probability that between 43% and 57% will be girls

We want to compute \( P(0.43 \times 200 \leq X_{\text{girls}} \leq 0.57 \times 200) \), which translates to:

$$P(86 \leq X_{\text{girls}} \leq 114)$$

Since \( X_{\text{girls}} = 200 - X_{\text{boys}} \), we are actually looking for:

$$P(86 \leq X_{\text{boys}} \leq 114)$$

To compute this, we standardize both limits:

For \( X = 86 \):

$$Z = \frac{86 - 100}{7.071} = \frac{-14}{7.071} \approx -1.98$$

For \( X = 114 \):

$$Z = \frac{114 - 100}{7.071} = \frac{14}{7.071} \approx 1.98$$

Thus, we compute the probability:

$$P(-1.98 < Z < 1.98) = P(Z < 1.98) - P(Z < -1.98)$$

From standard normal tables:

$$P(Z < 1.98) \approx 0.9761, \quad P(Z < -1.98) \approx 0.0239$$

So:

$$P(-1.98 < Z < 1.98) = 0.9761 - 0.0239 = 0.9522$$

Thus, the probability that between 43% and 57% will be girls is approximately:

$$P(86 \leq X_{\text{girls}} \leq 114) \approx 0.9522$$

---

### (c) Probability that more than 54% will be boys


We want to compute \( P(X > 0.54 \times 200) \), which translates to:


$$P(X > 108)$$


Standardize this value:


$$Z = \frac{108 - 100}{7.071} = \frac{8}{7.071} \approx 1.13$$


Now, using the cumulative distribution function:


$$P(Z > 1.13) = 1 - P(Z < 1.13)$$


From the standard normal table:


$$P(Z < 1.13) \approx 0.8708$$


So:


$$P(Z > 1.13) = 1 - 0.8708 = 0.1292$$


Thus, the probability that more than 54% will be boys is approximately:

$$P(X > 108) \approx 0.1292$$


---

### Summary of Results

- **(a)**: The probability that less than 40% of the next 200 children born will be boys is approximately \( 0.0023 \).
- **(b)**: The probability that between 43% and 57% will be girls is approximately \( 0.9522 \).
- **(c)**: The probability that more than 54% will be boys is approximately \( 0.1292 \).

## Problem 3

### 8.49 The credit hour distribution at Metropolitan Technological College is as follows:
```{r table, echo=FALSE}
x <- c(6, 9, 12, 15, 18)
p_x <- c(0.1, 0.2, 0.4, 0.2, 0.1)

data <- data.frame(x, p_x)

kable(data, col.names = c("x", "p(x)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

mu <- sum(x * p_x)

variance <- sum((x - mu)^2 * p_x)
```

Find \( \mu \) and \( \sigma^2 \). Give the 25 (with replacement) possible samples of size 2, their means, and their probabilities.


The mean is `r mu`

The mean \( \mu \) is calculated as:

$$\mu = \sum{x \cdot p(x)} = (6 \cdot 0.1) + (9 \cdot 0.2) + (12 \cdot 0.4) + (15 \cdot 0.2) + (18 \cdot 0.1)$$

$$\mu = 12$$

The variance is `r variance`

The variance \( \sigma^2 \) is calculated as:

$$
\sigma^2 = \sum{(x - \mu)^2 \cdot p(x)} = (6 - 12)^2 \cdot 0.1 + (9 - 12)^2 \cdot 0.2 + (12 - 12)^2 \cdot 0.4 + (15 - 12)^2 \cdot 0.2 + (18 - 12)^2 \cdot 0.1
$$

Substitute the differences:

$$
\sigma^2 = (-6)^2 \cdot 0.1 + (-3)^2 \cdot 0.2 + (0)^2 \cdot 0.4 + (3)^2 \cdot 0.2 + (6)^2 \cdot 0.1
$$

Now calculate the squares and multiply:

$$
\sigma^2 = 36 \cdot 0.1 + 9 \cdot 0.2 + 0 \cdot 0.4 + 9 \cdot 0.2 + 36 \cdot 0.1
$$

$$
\sigma^2 = 3.6 + 1.8 + 0 + 1.8 + 3.6
$$

Finally, sum the values:

$$
\sigma^2 = 10.8
$$


The 25 (with replacement) possible samples of size 2, their means, and their probabilities.

```{r table2, echo=FALSE}

samples <- expand.grid(x, x)

samples$Mean <- rowMeans(samples)

samples$Probability <- p_x[match(samples$Var1, x)] * p_x[match(samples$Var2, x)]

result <- samples[, c("Var1", "Var2", "Mean", "Probability")]
colnames(result) <- c("Sample 1", "Sample 2", "Mean", "Probability")

kable(result, digits = 4, caption = "Possible Samples of Size 2, Their Means, and Probabilities") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```